Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-07-17 draftfinalreport_v4.Rnw
add bit on why we use median instead of mean for future outbreak sizes
update captions to cost/benefit tables to indicate that we're using medians rather than means

The expected number of cases in New Zealand, assuming homogenous mixing, in a na\"{\i}ve population of subsequently lower (approximately 8\%) numbers and assuming measles $R_v$ is <1 is shown in \autoref{fig:sim} and \autoref{fig:sim1}. These simulations show that even in scenarios when $R_v$ is lower than one, and thus deterministically should fail to persist (i.e. become endemic or cause an outbreak), outbreaks can occur due to stochastic processes following measles importation. The most likely scenario is that very few cases occur, with the median value from 1000 simulations just 2 cases. Thus most measles introductions will be single cases or lead to single secondary cases. In this modelling exercise the mean value was 61 cases, suggesting there is likely to be an outbreak of a larger size given enough introductions. The maximum predicted number of cases was nearly 8,000 cases (\autoref{fig:sim1}). Note, however, that larger values are very rarely predicted and this does not take into account spatial variation or heterogeneous contact rates.

@@

\begin{sidewaystable}\small
<<results=tex,echo=FALSE>>=
colnames(costtable20)<-c("DHB", "Vacc", "Vacc costs", "Wages saved", "Manage saved", "Hosp saved", "Contacts","Costs saved", "Outbreak", "OB costs", "B/C")
costtable20[,2:10] <- round(costtable20[,2:10])
costtable20[,11] <- round(costtable20[,11],2)
latex(costtable20, file="", table.env=FALSE,rowname=NULL)
@
\caption{Benefit--cost analyses with 20 dollars per immunised person. \textit{Vacc} is numbers to immunise (see \autoref{table:attack}); \textit{Vacc costs} is cost for the catch up immunisation programme; \textit{Wages saved} is wages of care givers and cases saved (\$839 per case); \textit{Contacts} is costs saved through contact exclusion (\$2,624.57 per case); \textit{Manage saved} is management costs saved (\$1,765 per case), plus \$20 GP costs; \textit{Hosp saved} is the hospitalisation costs saved (\$1,877 per case); \textit{Costs save}  is the discounted costs saved; \textit{Outbreak}	is the predicted mean outbreak size despite $R_v < 1$ due to measles importation from 1000 simulations\footnote{Note the mean takes into account rare, but large outbreaks, see \autoref{fig:sim} and \autoref{fig:sim1}. Median values are all 2, thus the expected number of cases will be small. The mean is used as a conservative estimate regarding the benefits of immunisation, because it allows for rare but large outbreaks}; \textit{OB costs}	is costs expected due to continued measles importations based on 10 introductions of measles, one per year, but costs discounted on the same discounted rate; \textit{B/C} is the benefit--cost ratio.}
\label{table:cost20}
\end{sidewaystable}

\begin{sidewaystable}\small
<<results=tex,echo=FALSE>>=
colnames(costtable50)<-c("DHB", "Vacc", "Vacc costs", "Wages saved", "Manage saved", "Hosp saved", "Contacts","Costs saved", "Outbreak", "OB costs", "B/C")
costtable50[,2:10] <- round(costtable50[,2:10])
costtable50[,11] <- round(costtable50[,11],2)
latex(costtable50, file="", table.env=FALSE,rowname=NULL)
@
\caption{Benefit--cost analyses with 50 dollars per immunised person. \textit{Vacc} is numbers to immunise (see \autoref{table:attack}); \textit{Vacc costs} is cost for the catch up immunisation programme; \textit{Wages saved} is wages of care givers and cases saved (\$839 per case); \textit{Contacts} is costs saved through contact exclusion (\$2,624.57 per case); \textit{Manage saved} is management costs saved (\$1,765 per case), plus \$20 GP costs; \textit{Hosp saved} is the hospitalisation costs saved (\$1,877 per case); \textit{Costs save}  is the discounted costs saved; \textit{Outbreak}  is the predicted mean outbreak size despite $R_v < 1$ due to measles importation from 1000 simulations\footnote{Note the mean takes into account rare, but large outbreaks, see \autoref{fig:sim} and \autoref{fig:sim1}. Median values are all 2, thus the expected number of cases will be small. The mean is used as a conservative estimate regarding the benefits of immunisation, because it allows for rare but large outbreaks}; \textit{OB costs}	is costs expected due to continued measles importations based on 10 introductions of measles, one per year, but costs discounted on the same discounted rate; \textit{B/C} is the benefit--cost ratio.}
\label{table:cost50}
\end{sidewaystable}

>>>>

The expected number of cases in New Zealand, assuming homogenous mixing, in a na\"{\i}ve population of subsequently lower (approximately 8\%) numbers and assuming measles $R_v$ is <1 is shown in \autoref{fig:sim} and \autoref{fig:sim1}. These simulations show that even in scenarios when $R_v$ is lower than one, and thus deterministically should fail to persist (i.e. become endemic or cause an outbreak), outbreaks can occur due to stochastic processes following measles importation. The most likely scenario is that very few cases occur, with the median value from 1000 simulations just 2 cases. Thus most measles introductions will be single cases or lead to single secondary cases. In this modelling exercise the mean value was 61 cases, suggesting there is likely to be an outbreak of a larger size given enough introductions. The maximum predicted number of cases was nearly 8,000 cases (\autoref{fig:sim1}). Note, however, that larger values are very rarely predicted and this does not take into account spatial variation or heterogeneous contact rates. Because of this, we use the median values for predicting outbreak sizes per annum per DHB.

@@

\begin{sidewaystable}\small
<<results=tex,echo=FALSE>>=
colnames(costtable20)<-c("DHB", "Vacc", "Vacc costs", "Wages saved", "Manage saved", "Hosp saved", "Contacts","Costs saved", "Outbreak", "OB costs", "B/C")
costtable20[,2:10] <- round(costtable20[,2:10])
costtable20[,11] <- round(costtable20[,11],2)
latex(costtable20, file="", table.env=FALSE,rowname=NULL)
@
\caption{Benefit--cost analyses with 20 dollars per immunised person. \textit{Vacc} is numbers to immunise (see \autoref{table:attack}); \textit{Vacc costs} is cost for the catch up immunisation programme; \textit{Wages saved} is wages of care givers and cases saved (\$839 per case); \textit{Contacts} is costs saved through contact exclusion (\$2,624.57 per case); \textit{Manage saved} is management costs saved (\$1,765 per case), plus \$20 GP costs; \textit{Hosp saved} is the hospitalisation costs saved (\$1,877 per case); \textit{Costs save}  is the discounted costs saved; \textit{Outbreak}	is the predicted median outbreak size despite $R_v < 1$ due to measles importation from 1000 simulations\footnote{Note the median outbreak sizes are what is expected, while the mean takes into account rare, but large outbreaks, see \autoref{fig:sim} and \autoref{fig:sim1}.}; \textit{OB costs}	is costs expected due to continued measles importations based on 10 introductions of measles, one per year, but costs discounted on the same discounted rate; \textit{B/C} is the benefit--cost ratio.}
\label{table:cost20}
\end{sidewaystable}

\begin{sidewaystable}\small
<<results=tex,echo=FALSE>>=
colnames(costtable50)<-c("DHB", "Vacc", "Vacc costs", "Wages saved", "Manage saved", "Hosp saved", "Contacts","Costs saved", "Outbreak", "OB costs", "B/C")
costtable50[,2:10] <- round(costtable50[,2:10])
costtable50[,11] <- round(costtable50[,11],2)
latex(costtable50, file="", table.env=FALSE,rowname=NULL)
@
\caption{Benefit--cost analyses with 50 dollars per immunised person. \textit{Vacc} is numbers to immunise (see \autoref{table:attack}); \textit{Vacc costs} is cost for the catch up immunisation programme; \textit{Wages saved} is wages of care givers and cases saved (\$839 per case); \textit{Contacts} is costs saved through contact exclusion (\$2,624.57 per case); \textit{Manage saved} is management costs saved (\$1,765 per case), plus \$20 GP costs; \textit{Hosp saved} is the hospitalisation costs saved (\$1,877 per case); \textit{Costs save}  is the discounted costs saved; \textit{Outbreak}	is the predicted median outbreak size despite $R_v < 1$ due to measles importation from 1000 simulations\footnote{Note the median outbreak sizes are what is expected, while the mean takes into account rare, but large outbreaks, see \autoref{fig:sim} and \autoref{fig:sim1}.}; \textit{OB costs}	is costs expected due to continued measles importations based on 10 introductions of measles, one per year, but costs discounted on the same discounted rate; \textit{B/C} is the benefit--cost ratio.}
\label{table:cost50}
\end{sidewaystable}
